---
title: "BDA400 Assignment 2 - Technical Analysis using R (Preliminary Stage)"
author: "Gopi Sharan Gajmer Sunar"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.width = 9, fig.height = 4.5)
```

# 1. Packages

Run this chunk once to install anything missing, then load the libraries.

```{r packages}
required_packages <- c("quantmod", "TTR", "xts", "zoo", "ggplot2", "knitr")
missing <- required_packages[!(required_packages %in% installed.packages()[, "Package"])]
if (length(missing) > 0) install.packages(missing)

library(quantmod)
library(TTR)
library(ggplot2)
library(knitr)
```

# 2. Configuration

`portfolio.txt` lists one ticker per line (lines starting with `#` are ignored).

```{r config}
portfolio_file <- "portfolio.txt"   # keep this file in the same folder as this .Rmd
start_date     <- as.Date("2023-01-01")
end_date       <- Sys.Date()
ma_windows     <- c(20, 50)         # moving-average windows (trading days)
```

# 3. Utility functions

## 3.1 Read portfolio and import stock data

```{r load-functions}
read_portfolio <- function(path = "portfolio.txt") {
  if (!file.exists(path)) stop("Cannot find portfolio file: ", path)
  symbols <- trimws(readLines(path, warn = FALSE))
  symbols <- symbols[symbols != "" & !startsWith(symbols, "#")]
  unique(toupper(symbols))
}

# Reads portfolio.txt, downloads each symbol with quantmod, and returns a
# named list with ONE DATA FRAME PER STOCK SYMBOL.
load_stock_data <- function(path = "portfolio.txt", from, to = Sys.Date()) {
  symbols <- read_portfolio(path)
  stocks  <- list()

  for (sym in symbols) {
    xt <- tryCatch(
      getSymbols(sym, src = "yahoo", from = from, to = to, auto.assign = FALSE),
      error = function(e) {
        warning("Could not download ", sym, ": ", conditionMessage(e))
        NULL
      }
    )
    if (is.null(xt)) next

    df <- data.frame(Date = index(xt), coredata(xt))
    names(df) <- c("Date", "Open", "High", "Low", "Close", "Volume", "Adjusted")
    stocks[[sym]] <- df
  }
  stocks
}
```

## 3.2 Compute statistics

```{r stats-functions}
# Statistical mode (most frequent value); prices are rounded so ties are meaningful
get_mode <- function(x, digits = 0) {
  x  <- round(na.omit(x), digits)
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}

# Takes one stock's data frame; returns a list with
#   $summary  - one-row data frame of statistics
#   $data     - the original data frame plus moving-average columns
calculate_statistics <- function(df, price_col = "Close", windows = c(20, 50)) {
  prices <- df[[price_col]]

  # Moving averages (simple) using TTR::SMA
  for (w in windows) df[[paste0("SMA", w)]] <- SMA(prices, n = w)

  summary_df <- data.frame(
    Mean   = mean(prices, na.rm = TRUE),
    Median = median(prices, na.rm = TRUE),
    Mode   = get_mode(prices),
    StdDev = sd(prices, na.rm = TRUE),
    Min    = min(prices, na.rm = TRUE),
    Max    = max(prices, na.rm = TRUE)
  )
  for (w in windows) {
    ma_col <- df[[paste0("SMA", w)]]
    summary_df[[paste0("Latest_SMA", w)]] <- tail(na.omit(ma_col), 1)
  }
  list(summary = round(summary_df, 2), data = df)
}

# Runs calculate_statistics() on every stock in the list
calculate_all_statistics <- function(stocks, windows = c(20, 50)) {
  lapply(stocks, calculate_statistics, windows = windows)
}
```

## 3.3 Display utilities

```{r display-functions}
# Table view (like Yahoo Finance "Historical Data")
display_stock_table <- function(df, symbol, n = 10) {
  cat("\n\n### ", symbol, " - most recent ", n, " trading days\n\n", sep = "")
  print(kable(tail(df[, c("Date", "Open", "High", "Low", "Close", "Adjusted", "Volume")], n),
              digits = 2, row.names = FALSE))
}

# Line chart: close price + moving averages
plot_price_ma <- function(df, symbol, windows = c(20, 50)) {
  ma_cols <- paste0("SMA", windows)
  long <- do.call(rbind, lapply(c("Close", ma_cols), function(s)
    data.frame(Date = df$Date, Series = s, Price = df[[s]])))
  ggplot(long, aes(Date, Price, colour = Series)) +
    geom_line(na.rm = TRUE) +
    labs(title = paste(symbol, "- Close price and moving averages"), y = "Price (USD)") +
    theme_minimal()
}

# Histogram of closing prices with mean / median / mode marked
plot_distribution <- function(df, symbol, stats) {
  s <- stats$summary
  ref <- data.frame(Stat = c("Mean", "Median", "Mode"), Value = c(s$Mean, s$Median, s$Mode))
  ggplot(df, aes(Close)) +
    geom_histogram(bins = 30, fill = "steelblue", alpha = 0.6) +
    geom_vline(data = ref, aes(xintercept = Value, colour = Stat), linewidth = 1) +
    labs(title = paste(symbol, "- Distribution of closing prices"), x = "Close (USD)") +
    theme_minimal()
}

# Volume bar chart
plot_volume <- function(df, symbol) {
  ggplot(df, aes(Date, Volume)) +
    geom_col(fill = "grey40") +
    labs(title = paste(symbol, "- Trading volume")) +
    theme_minimal()
}

# Candlestick chart using quantmod (last `months` months)
plot_candlestick <- function(df, symbol, months = 6) {
  xt <- xts::xts(df[, c("Open", "High", "Low", "Close", "Volume", "Adjusted")], order.by = df$Date)
  chartSeries(xt, name = symbol, subset = paste("last", months, "months"),
              theme = chartTheme("white"), TA = "addSMA(20);addSMA(50);addVo()")
}
```

# 4. Run the analysis

```{r run}
stocks <- load_stock_data(portfolio_file, from = start_date, to = end_date)
stats  <- calculate_all_statistics(stocks, windows = ma_windows)
names(stocks)   # symbols successfully loaded
```

# 5. Summary statistics for all stocks

```{r summary-table}
summary_all <- do.call(rbind, lapply(names(stats), function(s)
  cbind(Symbol = s, stats[[s]]$summary)))
kable(summary_all, row.names = FALSE, caption = "Statistics on closing prices")
```

# 6. Per-stock data and charts

```{r per-stock, results='asis', fig.show='asis'}
for (sym in names(stocks)) {
  cat("\n\n## ", sym, "\n")
  display_stock_table(stocks[[sym]], sym)
  cat("\n\n")
  print(plot_price_ma(stats[[sym]]$data, sym, ma_windows))
  print(plot_distribution(stocks[[sym]], sym, stats[[sym]]))
  print(plot_volume(stocks[[sym]], sym))
}
```

# 7. Candlestick charts

```{r candles, fig.height=6}
for (sym in names(stocks)) plot_candlestick(stocks[[sym]], sym)
```
